---
title: "Quadmods Import Script"
author: "Kyle MacDonald"
date: "January 16, 2016"
output: html_document
---

This script imports and formats data for active/passive learning experiment
about quadrilaterals. 

# Libraries

```{r, message=FALSE}
rm(list=ls())
library(jsonlite)
library(magrittr)
library(plyr)
library(dplyr)
library(lubridate)
library(tidyr)
```

Set working directory to where you stored raw JSON files.

```{r}
setwd("../raw-data/pilot/production-results/")
```

# Read in all the JSON strings

```{r comment=NA}
read.file <- function(filename) {
  con <- file(filename, "r", blocking = TRUE)
  lines <- paste(readLines(con, warn = FALSE), collapse = "\n")
  close(con)
  lines
}

json.filenames <- list.files()
json.strings <- Map(read.file, json.filenames)
```

# Convert JSON to a R data frame

Merge all the json strings together, as if they are in a big array, and convert it to an R data frame:

```{r comment=NA}
json.superstring <- paste("[",paste(json.strings, collapse = ","),"]")
assignments <- fromJSON( json.superstring )
```

Only keep unique worker IDs.

```{r comment=NA}
assignments <- distinct(assignments, WorkerId)
```

## Build experiment information data frame.

```{r comment=NA}
exp_df <- data.frame(
    workerid = assignments$WorkerId, 
    age = assignments$answers$age,
    gender = assignments$answers$gender,
    comments = assignments$answers$comment,
    exp_about = assignments$answers$about,
    approval_time = assignments$AutoApprovalTime,
    exp_time = assignments$answers$exp_time,
    training_condition = assignments$answers$training_condition,
    shape_learned = assignments$answers$shape_of_focus,
    training_time = assignments$answers$training_time,
    stringsAsFactors = F
)

```

## Build trial level data frame

```{r}
trial_data <- Map(function(id, subject.trial.data) 
    { cbind(workerid = id, subject.trial.data) },
                  assignments$WorkerId,
                  assignments$answers$data)

```

```{r}
strip.rownames <- function(x) {
  rownames(x) <- NULL
  x
}

trial_data <- strip.rownames(do.call(rbind, trial_data))
```

## Merge

```{r}
df_final <- left_join(exp_df, trial_data, by = "workerid")
```

## Fix timestamps

```{r}
# This drops the weirdo T and Z from the turk timestamps and then parses them into seconds time 0 (some UTC standard)
df_final$approval_time <- parse_date_time(
    gsub("[[:alpha:]]"," ", 
         df_final$approval_time),
    "Ymd hms"
)
```

## Do some clean-up:

- Make all shape and question strings singular
- Make a variable that has both the question and the shape so we can easily see if the trial is correct

```{r}
df_final <- df_final %>% 
    mutate(shape = ifelse(shape == "rectangles", "rectangle", 
                          ifelse(shape == "rhombuses", "rhombus", 
                                 ifelse(shape == "squares", "square",
                                        ifelse(shape == "parallelograms", "parallelogram", shape)))),
           question = ifelse(question == "rectangles", "rectangle", 
                          ifelse(question == "rhombuses", "rhombus", 
                                 ifelse(question == "squares", "square",
                                        ifelse(question == "parallelograms", "parallelogram", 
                                               ifelse(question == "2", "rhombus", question))))),
           shape_learned = ifelse(shape_learned == 2, "rhombus", shape_learned),
           question_and_shape = ifelse(trial_type == "relational", paste(question, shape, sep = '_'),
                                       paste(shape, question, sep = '_'))
           )
```

## Fix posttest data storage

Some code to recover entity posttest responses from the pilot data because 
there was a bug in the javascript that screwed up the primary method of data storage.

Note: this only should be run this on the first batch of 40 ss in the pilot data.

```{r, eval = f}
e_trial_data <- Map(function(id, e_post_test_responses, shape, question, trial_type, block) 
    { cbind(workerid = id, e_post_test_responses, shape, question, trial_type, block) },
                  assignments$WorkerId,
                  assignments$answers$e_posttest_responses,
                  assignments$answers$e_posttest_shape_order_as_presented,
                  assignments$answers$e_posttest_q_order_as_presented,
                  "entity",
                  "posttest"
)

e_trial_data <- data.frame(strip.rownames(do.call(rbind, e_trial_data)))
e_trial_data %<>% mutate(response_good = ifelse(e_post_test_responses == T, "yes", "no"))

df_final <- left_join(df_final, e_trial_data)

df_final %<>% mutate(response = ifelse(trial_type == "entity" & block == "posttest",
                                           response_good, response)) %>% 
    select(-e_post_test_responses, -response_good)
```

## Flag if a trial is correct

```{r}
# flag correct key
flag_correct_key <- function (question_and_shape) {
    switch(question_and_shape,
           parallelogram_rectangle = {correct_answer <- "no"},
           rectangle_rhombus = {correct_answer <- "no"},
           rhombus_rectangle = {correct_answer <- "no"},
           rhombus_square = {correct_answer <- "no"},
           square_rhombus = {correct_answer <- "yes"},
           square_parallelogram = {correct_answer <- "yes"},
           rectangle_square = {correct_answer <- "no"},
           parallelogram_rhombus = {correct_answer <- "no"},
           parallelogram_square = {correct_answer <- "no"},
           square_rectangle = {correct_answer <- "yes"},
           rhombus_parallelogram = {correct_answer <- "yes"},
           rectangle_parallelogram = {correct_answer <- "yes"},
           parallelogram_parallelogram = {correct_answer <- "yes"},
           rhombus_rhombus = {correct_answer <- "yes"},
           square_square = {correct_answer <- "yes"},
           rectangle_rectangle = {correct_answer <- "yes"})
    
    return(correct_answer)
}

# takes in a row of a data frame
# and based on values in that row
# returns if the trial was correct or incorrect

flag_correct_trial <- function (trial) {
    # grab stuff we need from the trial vector
    question_and_shape <- as.character(trial[which(names(trial)=="question_and_shape")])
    response <- as.character(trial[which(names(trial)=="response")])
    # check answer key
    correct_answer <- flag_correct_key(question_and_shape)
    # check ss response against the key and return T/F for each trial
    correct <- response == correct_answer
    return(correct)
}

# create a vector of correct/incorrect
correct <- apply(df_final, 1, flag_correct_trial)

# merge correct info with main df
df_final <- cbind(df_final, correct)
```

Anonymize

```{r}
anonymized_df <- df_final %>% 
    select(workerid) %>% 
    distinct() %>% 
    mutate(subids = 1:nrow(.))

df_final_clean <- left_join(df_final, anonymized_df, by = "workerid") %>% 
    select(-workerid)

# move subid to the first col
df_final_clean <- select(df_final_clean, subids, everything())
```

## Save tidy data 

```{r}
write.csv(x = df_final_clean, file = "../../../data/quadmods-sequences-pilot.csv", row.names=F)
```
