---
title: "Quadmods Import Script"
author: "Kyle MacDonald"
date: "January 16, 2016"
output: html_document
---

This script imports and formats data for baseline passive learning experiment
about quadrilaterals. 

# Libraries

```{r, message=FALSE}
rm(list=ls())
library(jsonlite)
library(magrittr)
library(plyr)
library(dplyr)
library(lubridate)
library(tidyr)
```

Set working directory to where you stored raw JSON files.

```{r}
setwd("../../raw-data/quadmods_baseline_full/production-results//")
```

# Read in all the JSON strings

```{r comment=NA}
read.file <- function(filename) {
  con <- file(filename, "r", blocking = TRUE)
  lines <- paste(readLines(con, warn = FALSE), collapse = "\n")
  close(con)
  lines
}

json.filenames <- list.files()
json.strings <- Map(read.file, json.filenames)
```

# Convert JSON to a R data frame

Merge all the json strings together, as if they are in a big array, and convert it to an R data frame:

```{r comment=NA}
json.superstring <- paste("[",paste(json.strings, collapse = ","),"]")
assignments <- fromJSON( json.superstring )
```

Only keep unique worker IDs.

```{r comment=NA}
assignments <- distinct(assignments, WorkerId)
```

## Build experiment information data frame.

```{r comment=NA}
exp_df <- data.frame(
    workerid = assignments$WorkerId, 
    age = assignments$answers$age,
    gender = assignments$answers$gender,
    comments = assignments$answers$comment,
    exp_about = assignments$answers$about,
    exp_unclear = assignments$answers$unclear,
    exp_strategy = assignments$answers$strategy,
    exp_external_aid = assignments$answers$external_aid,
    exp_improve = assignments$answers$better,
    approval_time = assignments$AutoApprovalTime,
    exp_time = assignments$answers$exp_total_time,
    stringsAsFactors = F
)
```

## Build trial level data frame

```{r}
trial_data <- Map(function(id, subject.trial.data) 
    { cbind(workerid = id, subject.trial.data) },
                  assignments$WorkerId,
                  assignments$answers$data)

```

```{r}
strip.rownames <- function(x) {
  rownames(x) <- NULL
  x
}

trial_data <- strip.rownames(do.call(rbind, trial_data))
```

## Merge

```{r}
df_final <- left_join(exp_df, trial_data, by = "workerid")
```

## Fix timestamps

```{r}
# This drops the weirdo T and Z from the turk timestamps and then parses them into seconds time 0 (some UTC standard)
df_final$approval_time <- parse_date_time(
    gsub("[[:alpha:]]"," ", 
         df_final$approval_time),
    "Ymd hms"
)
```

## Anonymize

```{r}
anonymized_df <- df_final %>% 
    select(workerid) %>% 
    distinct() %>% 
    mutate(subids = 1:nrow(.))

df_final_clean <- left_join(df_final, anonymized_df, by = "workerid") %>% 
    select(-workerid)

# move subid to the first col and add a variable to track both question and shape
df_final_clean <- select(df_final_clean, subids, everything()) %>% 
    mutate(question_and_shape = paste(question, shape, sep = '_'))
```

## Save tidy data 

```{r}
write.csv(x = df_final_clean, file = "../../../data/baseline_expt/quadmods-baseline.csv",
          row.names=F)
```
