---
title: 'Quadrilateral learning experiments: Baseline'
output: pdf_document
date: "January 22, 2016"
---

```{r, echo = F}
rm(list=ls()) # clear workspace
knitr::opts_chunk$set(
  fig.height=4, 
  fig.width=7)
```

```{r, results=F}
library(langcog)
library(dplyr)
library(ggplot2)
library(tidyr)
library(binom)
library(lme4)
library(bootstrap)
library(magrittr)
library(stringr)
theme_set(theme_classic())
```

Load data.

```{r data_clean}
d1 <- read.csv("../../data/baseline_expt/quadmods-baseline.csv", header=TRUE, 
               row.names=NULL, stringsAsFactors = FALSE)

d1 <- d1 %>% 
    distinct() %>% 
    mutate(block_num = as.numeric(block))
```

Get the block each ss reached criterion

```{r}
d1 %<>% 
    filter(trial_type == "training") %>% 
    group_by(subids) %>% 
    summarise(block_reached_crit = max(block_num)) %>% 
    left_join(., d1) %>% 
    mutate(
      reached_crit = ifelse(block_reached_crit < 15, "yes", "no"),
      blocks_to_go = block_reached_crit - block_num + 1
      )
```

Descriptives
--------

### How many participants?

```{r}
d1 %>% 
    select(subids, reached_crit) %>% 
    unique() %>% 
    group_by(reached_crit) %>% 
    summarise(n_subs = n()) %>% 
    knitr::kable()
```

### How long did experiment take?

```{r}
d1 %>% 
    filter(trial_type == "training") %>%
    group_by(reached_crit) %>% 
    summarise(m_train_time_sec = (mean(trial_time) / 1000),
              m_exp_time_min = (mean(exp_time) / 1000) / 60) %>% 
    knitr::kable()

ss <- d1 %>% 
    filter(trial_type == "training") %>% 
    group_by(subids, reached_crit) %>% 
    summarise(train_time_sec = (mean(trial_time) / 1000),
              exp_time_min = (mean(exp_time) / 1000) / 60,
              max_training_block = max(as.numeric(block)))

ggplot(aes(x=exp_time_min, fill = reached_crit), data = ss) +
    geom_density(alpha = 0.5) +
    scale_fill_solarized() 
```

### How many blocks did it take to reach criterion? 

```{r}
ss <- d1 %>% 
    filter(trial_type == "training") %>% 
    group_by(subids) %>% 
    summarise(block_reached_crit = max(as.numeric(block)))

ggplot(data = ss, aes(x = block_reached_crit)) + 
    geom_histogram() +
    xlab("Block Reached Criterion") +
    theme(text = element_text(size = 20))
```

### Comments

```{r}
d1 %>% 
    distinct(subids) %>% 
    select(subids:exp_improve) %>% 
    knitr::kable()
```

Visualization
--------

### Training performance across blocks for all Ss aggregate

```{r}
ms <- d1 %>% 
  filter(trial_type == "training") %>%
  group_by(blocks_to_go) %>% 
  multi_boot_standard("correct") %>%
  mutate( 
    p_error = 1 - mean,
    ci_lower = 1 - ci_lower,
    ci_upper = 1 - ci_upper)
    

ggplot(data = ms, aes(x=as.numeric(blocks_to_go), y=p_error)) +
  geom_pointrange(aes(ymin=ci_lower, ymax=ci_upper)) +
  geom_line() +
  geom_hline(aes(yintercept = 0.5), linetype = "dashed") +
  ylim(0, .6) +
  xlim(15,0) + 
  scale_color_solarized() +
  ylab("Probability of error") +
  xlab("Blocks to go") +
  theme(text = element_text(size = 20))
```

### Training performance across blocks for all Ss by reached crit

```{r}
ms <- d1 %>% 
  filter(trial_type == "training") %>%
  group_by(blocks_to_go, reached_crit) %>% 
  multi_boot_standard("correct") %>%
  mutate( 
    p_error = 1 - mean,
    ci_lower = 1 - ci_lower,
    ci_upper = 1 - ci_upper)
    

ggplot(data = ms, aes(x=as.numeric(blocks_to_go), y=p_error, color=reached_crit)) +
  geom_pointrange(aes(ymin=ci_lower, ymax=ci_upper)) +
  geom_line() +
  geom_hline(aes(yintercept = 0.5), linetype = "dashed") +
  ylim(0, .6) +
  xlim(15,0) + 
  scale_color_solarized() +
  ylab("Probability of error") +
  xlab("Blocks to go") +
  theme(text = element_text(size = 20))
```

### Training performance across blocks for all Ss by shape

```{r}
ms <- d1 %>% 
  filter(trial_type == "training") %>%
  group_by(blocks_to_go, question) %>% 
  multi_boot_standard("correct") %>%
  mutate( 
    p_error = 1 - mean,
    ci_lower = 1 - ci_lower,
    ci_upper = 1 - ci_upper)
    

ggplot(data = ms, aes(x=blocks_to_go, y=p_error, color=c(question))) +
  geom_pointrange(aes(ymin=ci_lower, ymax=ci_upper)) +
  geom_line() +
  geom_hline(aes(yintercept = 0.5), linetype = "dashed") +
  ylim(0, .6) +
  xlim(15,0) + 
  scale_color_solarized() +
  ylab("Probability of error") +
  xlab("Blocks to go") +
  theme(text = element_text(size = 20))
```

### Training performance across blocks for all Ss by shape for those who reached criterion

```{r}
ms <- d1 %>% 
  filter(trial_type == "training", reached_crit=="yes") %>%
  group_by(blocks_to_go, question) %>% 
  multi_boot_standard("correct") %>%
  mutate( 
    p_error = 1 - mean,
    ci_lower = 1 - ci_lower,
    ci_upper = 1 - ci_upper)
    

ggplot(data = ms, aes(x=blocks_to_go, y=p_error, color=c(question))) +
  geom_pointrange(aes(ymin=ci_lower, ymax=ci_upper)) +
  geom_line() +
  geom_hline(aes(yintercept = 0.5), linetype = "dashed") +
  ylim(0, .6) +
  xlim(15,0) + 
  scale_color_solarized() +
  ylab("Probability of error") +
  xlab("Blocks to go") +
  ggtitle("Shapes for reached crit") +
  theme(text = element_text(size = 20))
```

### Training performance across blocks for all Ss by shape for those who did not reach criterion

```{r}
ms <- d1 %>% 
  filter(trial_type == "training", reached_crit=="no") %>%
  group_by(blocks_to_go, question) %>% 
  multi_boot_standard("correct") %>%
  mutate( 
    p_error = 1 - mean,
    ci_lower = 1 - ci_lower,
    ci_upper = 1 - ci_upper)
    

ggplot(data = ms, aes(x=blocks_to_go, y=p_error, color=c(question))) +
  geom_pointrange(aes(ymin=ci_lower, ymax=ci_upper)) +
  geom_line() +
  geom_hline(aes(yintercept = 0.5), linetype = "dashed") +
  ylim(0, .6) +
  xlim(15,0) + 
  scale_color_solarized() +
  ylab("Probability of error") +
  xlab("Blocks to go") +
  ggtitle("Shapes for did not reach crit") +
  theme(text = element_text(size = 20))
```

### Training performance across blocks split by ss who reached training criterion

```{r}
ms <- d1 %>% 
    filter(trial_type == "training") %>% 
    group_by(block, reached_crit) %>% 
    multi_boot_standard(column = "correct") %>% 
    mutate(prob_error = 1 - mean)

ggplot(data = ms, aes(x = as.numeric(block), y = mean, color = reached_crit)) + 
    geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
    geom_line() +
    geom_hline(aes(yintercept = 0.5), linetype = "dashed") +
    ylim(0.2, 1) +
    scale_color_solarized() +
    xlab("Block") +
    theme(text = element_text(size = 20))
```

### Average blocks to criterion

```{r}
d1 %>% 
    filter(trial_type == "training") %>% 
    mutate(condition = "baseline") %>% 
    group_by(subids, condition) %>% 
    summarise(block_reached_crit = max(as.numeric(block))) %>% 
    filter(block_reached_crit != 15) %>% # remove ss who didn't reach criterion
    group_by(condition) %>% 
    multi_boot_standard(column = "block_reached_crit") %>% 
    knitr::kable()
```

### Relational test broken down by question and block

```{r, fig.width=9, fig.height=6}
ms <- d1 %>% 
    filter(trial_type %in% c("relational")) %>%
    group_by(question_and_shape, question, block = as.factor(block), reached_crit) %>%
    multi_boot_standard(column = "correct") %>% 
    ungroup() %>% 
    mutate(block = relevel(block, "pretest")) 

ggplot(aes(x = question_and_shape, y = mean, fill = block), data = ms) + 
    geom_bar(stat="identity", position="dodge", width = 0.5) +
    geom_linerange(aes( ymin = ci_lower, ymax = ci_upper),  
                   position = position_dodge(width = 0.5)) +
    geom_hline(yintercept = .5, lty = 2) +
    ylim(0, 1) + 
    xlab("Question") +
    scale_fill_solarized() +
    facet_grid(reached_crit~.) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 30, hjust = 1),
          text = element_text(size = 20)) +
    ggtitle("Are all _____ also ______?") 
```

### Overall accuracy analysis for relational test across blocks

```{r}
ms <- d1 %>% 
    filter(trial_type %in% c("relational")) %>% 
    group_by(trial_type, block = as.factor(block)) %>%
    multi_boot_standard(column = "correct")

ms %<>% mutate(chance_line = ifelse(trial_type == "relational", 0.5, 0.5),
               block = relevel(block, "pretest"))

ggplot(aes(x = block, y = mean), data = ms) +
    geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
    geom_line(aes(group=1)) +
    geom_hline(aes(yintercept = chance_line), linetype = "dashed") +
    ylim(0,1.0) +
    scale_color_solarized() 
```

### Within subjects change scores for all shapes

```{r}
ss_acc <- d1 %>% 
    filter(trial_type %in% c("relational")) %>% 
    group_by(subids, block, trial_type, reached_crit) %>% 
    summarise(m_acc = mean(correct)) %>% 
    spread(key = block, value = m_acc) %>% 
    mutate(m_diff_score = posttest - pretest)

ms_change <- ss_acc %>% 
    group_by(trial_type) %>% 
    multi_boot_standard(column = "m_diff_score")

ggplot(aes(x = trial_type, y = mean, fill = trial_type), 
       data = ms_change) +
    geom_bar(stat = "identity", width = 0.5) +
    geom_linerange(aes(ymin = ci_lower, ymax = ci_upper)) +
    geom_hline(yintercept = 0) +
    scale_fill_solarized() +
    ylim(-0.2, 0.2) + 
    guides(fill=F)
```

## Analysis splitting by ss who reached criterion

### Overall accuracy analysis

```{r}
ms <- d1 %>% 
    filter(trial_type %in% c("relational")) %>% 
    group_by(trial_type, block = as.factor(block), reached_crit) %>%
    multi_boot_standard(column = "correct")


ms %<>% ungroup() %>%  mutate(block = relevel(block, "pretest"))

ggplot(aes(x = block, y = mean), data = ms) +
    geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
    geom_line(aes(group=1)) +
    geom_hline(aes(yintercept = 0.5), linetype = "dashed") +
    ylim(0,1.0) +
    facet_grid(.~reached_crit) +
    scale_color_solarized() +
    theme_bw() +
    theme(text = element_text(size = 20))
```

### Within subjects change scores

Distribution of change scores split by if ss reached criterion.

```{r}
ggplot(data = ss_acc, aes(x = m_diff_score, fill = reached_crit)) + 
    geom_density(alpha = 0.5) +
    scale_fill_solarized()
```

```{r}
ss_acc <- d1 %>% 
    filter(trial_type %in% c("relational")) %>% 
    group_by(subids, block, trial_type, reached_crit) %>% 
    summarise(m_acc = mean(correct)) %>% 
    spread(key = block, value = m_acc) %>% 
    mutate(m_diff_score = posttest - pretest)

# filter out ss 3SD away from mean
ss_acc %<>%
    mutate(include = ifelse(abs(m_diff_score) > sd(ss_acc$m_diff_score) * 3 +
                                mean(ss_acc$m_diff_score),
                            "no", "yes"))

ms_change <- ss_acc %>% 
    filter(include == "yes") %>% 
    group_by(trial_type, reached_crit) %>% 
    multi_boot_standard(column = "m_diff_score")

ggplot(aes(x = trial_type, y = mean, fill = trial_type), 
       data = ms_change) +
    geom_bar(stat = "identity", width = 0.5) +
    geom_linerange(aes(ymin = ci_lower, ymax = ci_upper)) +
    facet_grid(.~reached_crit) +
    geom_hline(yintercept = 0) +
    scale_fill_solarized() +
    ylim(-0.2, 0.2) + 
    guides(fill=F) + 
    theme_bw()
```
