---
title: "Quadrilateral learning experiments: Sequences"
author: "Kyle MacDonald and Michael Frank"
date: "January 22, 2016"
output: html_document
---

```{r, echo = F}
rm(list=ls()) # clear workspace
knitr::opts_chunk$set(warning=FALSE, message=FALSE, sanitize = T, 
                      fig.height=4, fig.width=7, echo=F, cache = T)
```

```{r}
library(langcog)
library(dplyr)
library(ggplot2)
library(tidyr)
library(binom)
library(lme4)
library(bootstrap)
library(magrittr)
library(stringr)
theme_set(theme_bw())

## for bootstrapping 95% confidence intervals
theta <- function(x,xdata,na.rm=T) {mean(xdata[x],na.rm=na.rm)}
ci.low <- function(x,na.rm=T) {
  mean(x,na.rm=na.rm) - quantile(bootstrap(1:length(x),1000,theta,x,na.rm=na.rm)$thetastar,.025,na.rm=na.rm)}
ci.high <- function(x,na.rm=T) {
  quantile(bootstrap(1:length(x),1000,theta,x,na.rm=na.rm)$thetastar,.975,na.rm=na.rm) - mean(x,na.rm=na.rm)}
```

Load data.

```{r data_clean}
d1 <- read.csv("../data/quadmods-sequences-exp1_2.csv", header=TRUE, 
               row.names=NULL, stringsAsFactors = FALSE)

d1$block <- factor(d1$block, levels = c("pretest", "posttest"))
d1 <- d1 %>% 
    filter(training_condition %in% c("active_active", "passive_passive")) %>% 
    distinct()
```

Descriptives
--------

### How many participants in each condition?

```{r}
d1 %>% 
    select(subids, training_condition, experiment) %>% 
    unique() %>% 
    group_by(training_condition, experiment) %>% 
    summarise(n())
```

### How long did each condition take?

```{r}
d1 %>% 
    group_by(training_condition, experiment) %>% 
    summarise(m_train_time = (mean(training_time) / 1000) / 60,
              m_exp_time = (mean(exp_time) / 1000) / 60)

ss <- d1 %>% 
    distinct(subids) %>% 
    mutate(train_time_min = (training_time / 1000) / 60,
              exp_time_min = (exp_time / 1000) / 60)

ggplot(aes(x=train_time_min), data = ss) +
    geom_histogram(binwidth = 0.1) +
    facet_grid(.~training_condition) 
```

### SS level filtering

Remove:
- Participants who were scored perfect on the pretest
- Participants who's difference scores were 3 SD away from the mean difference score

```{r}
ss_diff_scores <- d1 %>% 
    filter(trial_type %in% c("relational","entity")) %>% 
    group_by(subids, training_condition, block, trial_type) %>% 
    summarise(m = mean(correct)) %>% 
    spread(key = block, value = m) %>% 
    mutate(tot_diff_score = posttest - pretest)

ms <- ss_diff_scores %>% 
    group_by(training_condition, trial_type) %>% 
    summarise(stdev_diff = sd(tot_diff_score),
              mean_diff = mean(tot_diff_score),
              mean_posttest = mean(posttest),
              stdev_posttest = sd(posttest))

ss_diff_scores <- left_join(ss_diff_scores, ms)

# create filter based on sd from mean difference score
filter_ss <- function (diff_score, stdev_diff, mean_diff) {
    if (diff_score >= mean_diff + (2*stdev_diff) | diff_score <= mean_diff - (2*stdev_diff)) {
        remove <- "yes"
    } else {
        remove <- "no"
    }
    remove
}

ss_diff_scores <- ss_diff_scores %>% 
    rowwise() %>% 
    mutate(remove = filter_ss(diff_score = posttest, stdev_diff = stdev_posttest, mean_diff = mean_posttest))
```

```{r}
# merge filter info with rest of data
d1 <- ss_diff_scores %>% 
    select(subids, remove, trial_type) %>%
    distinct() %>% 
    left_join(d1, .)

d1 <- filter(d1, remove == "no")
```


Visualization
--------

### Relational test broken down by question and block

```{r, fig.width=9, fig.height=6}
ms <- d1 %>% 
  filter(trial_type %in% c("relational")) %>%
  group_by(question_and_shape, question, block) %>%
  summarise(m.cih = ci.high(correct),
            m.cil = ci.low(correct),
            m = mean(correct))

ggplot(aes(x = question_and_shape, y = m, fill = block), data = ms) + 
    geom_bar(stat="identity", position="dodge") +
    geom_linerange(aes( ymin = m - m.cil, ymax = m + m.cih),  position = position_dodge(width = 0.9)) +
    geom_hline(yintercept = .5, lty = 2) +
    ylim(0, 1) + 
    scale_fill_solarized() +
    theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
    ggtitle("Accuracy on each question: Relational Test")
```

### Entity tests broken down by question and block

```{r, fig.width = 9, fig.height=6}
ms <- d1 %>% 
  filter(trial_type %in% c("entity")) %>%
  group_by(question_and_shape, question, block) %>%
  summarise(m.cih = ci.high(correct),
            m.cil = ci.low(correct),
            m = mean(correct))

ggplot(aes(x = question_and_shape, y = m, fill = block), data = ms) + 
    geom_bar(stat="identity", position="dodge") +
    geom_linerange(aes( ymin = m - m.cil, ymax = m + m.cih),  position = position_dodge(width = 0.9)) +
    ylim(0, 1) + 
    scale_fill_solarized() +
    theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
    ggtitle("Accuracy on each question: Entity Test")
```

## Overall accuracy analysis for each test

```{r}
ms <- d1 %>% 
    filter(trial_type %in% c("relational", "entity")) %>% 
    group_by(trial_type, block = as.factor(block)) %>%
    summarise(m_acc = mean(correct),
              m.cih = ci.high(correct),
              m.cil = ci.low(correct))

ms %<>% mutate(chance_line = ifelse(trial_type == "relational", 0.5, 0.5))

ggplot(aes(x = block, y = m_acc), data = ms) +
    geom_pointrange(aes(ymin = m_acc - m.cil, ymax = m_acc + m.cih)) +
    geom_line(aes(group=1)) +
    geom_hline(aes(yintercept = chance_line), linetype = "dashed") +
    ylim(0.4,1.0) +
    facet_grid(.~trial_type) +
    scale_color_solarized() 
```

### Within subjects change scores for all shapes

```{r}
ss_acc <- d1 %>% 
    filter(trial_type %in% c("relational", "entity")) %>% 
    group_by(subids, block, trial_type) %>% 
    summarise(m_acc = mean(correct)) %>% 
    spread(key = block, value = m_acc) %>% 
    mutate(m_diff_score = posttest - pretest)

ms_change <- ss_acc %>% 
    group_by(trial_type) %>% 
    summarise(m_change_score = mean(m_diff_score),
              m.cih = ci.high(m_diff_score),
              m.cil = ci.low(m_diff_score))

ggplot(aes(x = trial_type, y = m_change_score, fill = trial_type), 
       data = ms_change) +
    geom_bar(stat = "identity") +
    geom_linerange(aes(ymin = m_change_score - m.cil, ymax = m_change_score + m.cih)) +
    geom_hline(yintercept = 0) +
    scale_fill_solarized() +
    ylim(-0.1, 0.1) + 
    guides(fill=F)
```

## Accuracy on the learned shape

```{r}
d_shape_learned <- filter(d1, question == shape_learned | shape == shape_learned)
```

### Relational tests for learned shape

```{r, fig.width=9, fig.height=6}
ms <- d_shape_learned %>% 
  filter(trial_type %in% c("relational")) %>%
  group_by(question_and_shape, question, block) %>%
  summarise(m.cih = ci.high(correct),
            m.cil = ci.low(correct),
            m = mean(correct))

ggplot(aes(x = question_and_shape, y = m, fill = block), data = ms) + 
    geom_bar(stat="identity", position="dodge") +
    geom_linerange(aes( ymin = m - m.cil, ymax = m + m.cih),  position = position_dodge(width = 0.9)) +
    geom_hline(yintercept = .5, lty = 2) +
    ylim(0, 1) + 
    scale_fill_solarized() +
    theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
    ggtitle("Rhombus: Relational Test")
```

### Entity tests for learned shape

```{r, fig.width = 9, fig.height=6}
ms <- d_shape_learned %>% 
  filter(trial_type %in% c("entity")) %>%
  group_by(question_and_shape, question, block) %>%
  summarise(m.cih = ci.high(correct),
            m.cil = ci.low(correct),
            m = mean(correct))

ggplot(aes(x = question_and_shape, y = m, fill = block), data = ms) + 
    geom_bar(stat="identity", position="dodge") +
    geom_linerange(aes( ymin = m - m.cil, ymax = m + m.cih),  position = position_dodge(width = 0.9)) +
    ylim(0, 1) + 
    scale_fill_solarized() +
    theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
    ggtitle("Rhombus: Entity Test")
```

### Overall accuracy change on Rhombus

```{r}
ms_shape <- d_shape_learned %>% 
    filter(trial_type %in% c("relational", "entity")) %>% 
    group_by(trial_type, block = as.factor(block)) %>%
    summarise(m_acc = mean(correct),
              m.cih = ci.high(correct),
              m.cil = ci.low(correct))

ms_shape$block <- factor(ms_shape$block, levels = c("pretest", "posttest"))
ms_shape %<>% mutate(chance_line = ifelse(trial_type == "relational", 0.5, 0.5))

ggplot(aes(x = block, y = m_acc), data = ms_shape) +
    geom_pointrange(aes(ymin = m_acc - m.cil, ymax = m_acc + m.cih)) +
    geom_line(aes(group=1)) +
    geom_hline(aes(yintercept = chance_line), linetype = "dashed") +
    ylim(0.4,1.0) +
    facet_grid(.~trial_type) +
    scale_color_solarized()
```

### Within subjects change scores for Rhombus

```{r}
ss_acc_learned <- d_shape_learned %>% 
    filter(trial_type %in% c("relational", "entity")) %>% 
    group_by(subids, block, trial_type) %>% 
    summarise(m_acc = mean(correct)) %>% 
    spread(key = block, value = m_acc) %>% 
    mutate(m_diff_score = posttest - pretest)

ms_change_learned <- ss_acc_learned %>% 
    group_by(trial_type) %>% 
    summarise(m_change_score = mean(m_diff_score),
              m.cih = ci.high(m_diff_score),
              m.cil = ci.low(m_diff_score))

ggplot(aes(x = trial_type, y = m_change_score, fill = trial_type), 
       data = ms_change_learned) +
    geom_bar(stat = "identity") +
    geom_linerange(aes(ymin = m_change_score - m.cil, ymax = m_change_score + m.cih)) +
    geom_hline(yintercept = 0) +
    scale_fill_solarized() +
    ylim(-0.2, 0.2) + 
    guides(fill=F)
```

Exploratory Analyses
--------------------

### Overall accuracy analysis collapsing across entity and relational tests shape learned

```{r}
ms <- d1 %>% 
    filter(trial_type %in% c("relational", "entity")) %>% 
    group_by(training_condition, block = as.factor(block)) %>% 
     summarise(m_acc = mean(correct),
              m.cih = ci.high(correct),
              m.cil = ci.low(correct))

ggplot(aes(x = block, y = m_acc, color = training_condition), data = ms) +
    geom_pointrange(aes(ymin = m_acc - m.cil, ymax = m_acc + m.cih)) +
    geom_line(aes(group=training_condition)) +
    ylim(0.4,0.9) +
    scale_color_solarized() 
```

### Overall accuracy for shape learned collapsing across tests

```{r}
ms <- d_shape_learned %>% 
    filter(trial_type %in% c("relational", "entity")) %>% 
    group_by(training_condition, block = as.factor(block)) %>% 
     summarise(m_acc = mean(correct),
              m.cih = ci.high(correct),
              m.cil = ci.low(correct))

ggplot(aes(x = block, y = m_acc, color = training_condition), data = ms) +
    geom_pointrange(aes(ymin = m_acc - m.cil, ymax = m_acc + m.cih)) +
    geom_line(aes(group=training_condition)) +
    ylim(0.4,0.9) +
    scale_color_solarized() 
```

### Total difference score analysis

```{r}
ss <- d1 %>% 
    filter(trial_type %in% c("relational","entity")) %>% 
    group_by(subids, training_condition, block, experiment) %>% 
    summarise(m = mean(correct)) %>% 
    spread(key = block, value = m) %>% 
    mutate(tot_diff_score = posttest - pretest)

ms <- ss %>% 
    group_by(training_condition, experiment) %>% 
    summarise(m = mean(tot_diff_score),
              ci_h = ci.high(tot_diff_score),
              ci_l = ci.low(tot_diff_score))
```

```{r}
ggplot(aes(x = training_condition, y = m, fill = training_condition), data = ms) +
    geom_bar(stat = "identity") +
    geom_linerange(aes(ymin = m - ci_l, ymax = m + ci_h)) +
    geom_hline(yintercept = 0) +
    guides(fill=F) +
    scale_fill_solarized() +
    facet_grid(.~experiment) +
    ylim(-0.2, 0.2)
```

### Total difference score analysis for learned shape

```{r}
ss <- d_shape_learned %>% 
    filter(trial_type %in% c("relational","entity")) %>% 
    group_by(subids, training_condition, block, experiment) %>% 
    summarise(m = mean(correct)) %>% 
    spread(key = block, value = m) %>% 
    mutate(tot_diff_score = posttest - pretest)

ms <- ss %>% 
    group_by(training_condition, experiment) %>% 
    summarise(m = mean(tot_diff_score),
              ci_h = ci.high(tot_diff_score),
              ci_l = ci.low(tot_diff_score))
```

Plot

```{r}
ggplot(aes(x = training_condition, y = m, fill = training_condition), data = ms) +
    geom_bar(stat = "identity") +
    geom_linerange(aes(ymin = m - ci_l, ymax = m + ci_h)) +
    geom_hline(yintercept = 0) +
    guides(fill=F) +
    scale_fill_solarized() +
    facet_grid(.~experiment) +
    ylim(-0.2, 0.2) +
    ggtitle("Total Difference Score for Learned Shape")
```

### Separate by question

```{r}
ss <- d1 %>% 
    filter(trial_type %in% c("relational","entity")) %>% 
    group_by(subids, training_condition, block, question, trial_type, experiment) %>% 
    summarise(m = mean(correct)) %>% 
    spread(key = block, value = m) %>% 
    mutate(tot_diff_score = posttest - pretest)

ms <- ss %>% 
    group_by(training_condition, question, trial_type, experiment) %>% 
    summarise(m = mean(tot_diff_score),
              ci_h = ci.high(tot_diff_score),
              ci_l = ci.low(tot_diff_score))

```

```{r}
ggplot(aes(x = question, y = m, fill = experiment), data = ms) +
    geom_bar(stat = "identity", position = "dodge") +
    geom_linerange(aes(ymin = m - ci_l, ymax = m + ci_h), position = position_dodge(width=0.9)) +
    geom_hline(yintercept = 0) +
    facet_grid(trial_type~training_condition) +
    scale_fill_solarized() +
    ylim(-0.25, 0.25) +
    ggtitle("Is a Rhombus also a ________?")
```

### Separate by shape tested

```{r}
ss <- d1 %>% 
    filter(trial_type %in% c("relational","entity")) %>% 
    group_by(subids, training_condition, block, shape, trial_type, experiment) %>% 
    summarise(m = mean(correct)) %>% 
    spread(key = block, value = m) %>% 
    mutate(tot_diff_score = posttest - pretest)

ms <- ss %>% 
    group_by(training_condition, shape, trial_type, experiment) %>% 
    summarise(m = mean(tot_diff_score),
              ci_h = ci.high(tot_diff_score),
              ci_l = ci.low(tot_diff_score))

```

```{r}
ggplot(aes(x = shape, y = m, fill = experiment), data = ms) +
    geom_bar(stat = "identity", position = "dodge") +
    geom_linerange(aes(ymin = m - ci_l, ymax = m + ci_h), position = position_dodge(width=0.9)) +
    geom_hline(yintercept = 0) +
    facet_grid(trial_type~training_condition) +
    scale_fill_solarized() +
    ylim(-0.3, 0.3) +
    ggtitle("Click on all of the Rhombuses")

```

### Analysis based on individual ss pretest knowledge of rhombus (entity)

Does your pretest rhombus knowledge predict how you will do at test? 
Do we see condition differences based on pretest rhombus knowledge

```{r}
ss_rh_pretest <- d1 %>% 
    filter(trial_type == "entity", shape == shape_learned | question == shape_learned, 
           block == "pretest") %>% 
    distinct() %>% 
    group_by(subids) %>% 
    summarise(num_corr_rh_pre = sum(correct)) %>% 
    mutate(num_corr_rh_pre_fact = cut(num_corr_rh_pre, 4))

# merge with full data frame
d1 <- left_join(d1, ss_rh_pretest, by = "subids")
```

```{r}
ss <- d1 %>% 
    filter(trial_type %in% c("relational","entity")) %>% 
    group_by(subids, training_condition, block, trial_type, num_corr_rh_pre_fact, num_corr_rh_pre) %>% 
    summarise(m = mean(correct)) %>% 
    spread(key = block, value = m) %>% 
    mutate(tot_diff_score = posttest - pretest)
```

```{r}
ms <- ss %>% 
    group_by(training_condition, num_corr_rh_pre_fact) %>% 
    summarise(m = mean(tot_diff_score),
              ci_h = ci.high(tot_diff_score),
              ci_l = ci.low(tot_diff_score))


ggplot(aes(x = num_corr_rh_pre_fact, y = m, fill = training_condition), data = ms) +
    geom_bar(stat = "identity") +
    geom_linerange(aes(ymin = m - ci_l, ymax = m + ci_h)) +
    geom_hline(yintercept = 0) +
    facet_grid(.~training_condition) +
    guides(fill=F) +
    scale_fill_solarized() +
    ylim(-0.55, 0.55) 
```

```{r}
ggplot(aes(x = num_corr_rh_pre_fact, y = tot_diff_score, fill = training_condition), data = ss) +
    geom_boxplot() +
    scale_fill_solarized() +
    facet_grid(.~training_condition)
```

### Analysis based on evidence given or selected

Munge training data, so we can score them.

```{r, eval = F}
ss_sampling <- d1 %>% 
    filter(trial_type == "training", experiment == "random_passive", 
           training_condition %in% c("active_active", "passive_passive")) %>% 
    mutate(question_and_shape = paste(shape, question, sep = "_"))

correct <- apply(ss_sampling, 1, flag_correct_trial)
ss_sampling$correct <- NULL

# merge correct info with main df
ss_sampling <- cbind(ss_sampling, correct)
```

Score the training data for each participant

```{r}

```

Models (todo)
-------

### Predict overall accuracy based on condition and test type

```{r, eval = F}
mm1 <- glmer(correct ~ training_condition * block + (1|subids),
             data = filter(d1, trial_type %in% c("relational", "entity")),
             family=binomial, 
             nAGQ = 1,
             control = glmerControl(optimizer = "bobyqa"))
summary(mm1)
```

### Predict accuracy for learned shape based on condition and test type

```{r, eval = F}
mm2 <- glmer(correct ~ training_condition * block * trial_type + (1|subids) ,
             data = filter(d_shape_learned, trial_type %in% c("relational", "entity")),
             family=binomial, 
             nAGQ = 1,
             control = glmerControl(optimizer = "bobyqa"))
summary(mm2)
```

### Predict change scores for learned shape based on condition and test type

```{r, eval = F}
lm1 <- lm(m_diff_score ~ training_condition * trial_type,data = ss_acc)
summary(lm1)
```